---
title: "`tidyverse` - `ggplot2`"
author: "Pablo Benavides-Herrera"
date: '2020-06-02'
output:
  html_notebook:
    toc: yes
    toc_float: yes
    theme: united
  github_document:
    toc: yes
    dev: jpeg
subtitle: Lecture 3
---

# Plots


```{r pkgs, message=FALSE}
library(tidyverse)
library(plotly)
library(gridExtra)
library(patchwork)
library(nycflights13)
library(lubridate)
```


We will continue using the `mpg` dataset.

```{r mpg}
glimpse(mpg)
```

## Histograms

We start by plotting a simple histogram of the engine displacement.

```{r hist}
g <- ggplot(mpg, aes(displ)) + 
  scale_fill_brewer(palette = "Spectral")

g + geom_histogram(binwidth = .1, 
                   col="black", 
                   size=.1) +  # change binwidth
  labs(title="Histogram with Auto Binning", 
       subtitle="Engine Displacement across Vehicle Classes") 
```

We can create a stacked histogram by defining a fill aesthetics.

```{r stacked hist}
g + geom_histogram(aes(fill = class) ,
                   binwidth = .1, 
                   col="black", 
                   size=.1) +  # change binwidth
  labs(title="Histogram with Auto Binning", 
       subtitle="Engine Displacement across Vehicle Classes") 
```

Changing the number of bins can show different patterns. Setting the number of bins to 10:

```{r}
g +geom_histogram(aes(fill=class),
                   bins=10,
                   col="black", size = 0.1) +   # change number of bins
  labs(title="Histogram with Fixed Bins",
       subtitle="Engine Displacement across Vehicle Classes")
```


## Bar plots

If we want to plot factors, a **bar plot** could be a good way to do it.

```{r}
g <- ggplot(mpg, aes(manufacturer))
g + geom_bar(aes(fill=class), width = 0.5) +
  theme(axis.text.x = element_text(angle=65, vjust=0.7)) +
  labs(title="Histogram on Categorical Variable",
       subtitle="Manufacturer across Vehicle Classes")
```

## Density plots

```{r}
g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill=factor(cyl)), 
                 alpha=0.5) +
    labs(title="Density plot",
         subtitle="City Mileage Grouped by Number of cylinders",
         caption="Source: mpg",
         x="City Mileage",
         fill="# Cylinders")
```

## Boxplots


```{r}
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(varwidth=T, fill="dodgerblue", alpha = 0.5) +
    labs(title="Box plot",
         subtitle="City Mileage grouped by Class of vehicle",
         caption="Source: mpg",
         x="Class of Vehicle",
         y="City Mileage")
```

Setting `varwidth=T` adjusts the width of the boxes to be proportional to the number of observation it contains.
 
 
```{r}
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(aes(fill=factor(cyl))) +  scale_fill_brewer(palette = "Spectral") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  labs(title="Box plot",
       subtitle="City Mileage grouped by Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```
 
 
## Violin plot

```{r}
g <- ggplot(mpg, aes(class, cty))
p <- g + geom_violin() +  scale_fill_brewer(palette = "Spectral") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  labs(title="Violin plot",
       subtitle="City Mileage grouped by Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
p
```


```{r}
g + geom_violin(trim = T) + coord_flip()
```



```{r}
p + stat_summary(fun=median, geom="point", size=2, color="red")
```


```{r}
p + geom_boxplot(width = 0.1)
```


```{r}
g <- ggplot(mpg, aes(class, cty, color = class))
pc <- g + geom_violin() +
  labs(title="Violin plot",
       subtitle="City Mileage vs Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage") +
  theme(legend.position = "none")
pc
```



```{r}
pc + scale_color_brewer(palette = "Dark2")
```

## Arranging many plots 

### with `gridExtra`

```{r}
p_data <- ggplot(mpg, aes(x = hwy))
p1 <- p_data + geom_histogram(binwidth = 1, color = "grey30", fill = "white") + 
  labs(title="Bin width 1")
p2 <-  p_data + geom_histogram(binwidth = 2, color = "dodgerblue", 
                               fill = "blue", alpha = 0.4) + 
  labs(title="Bin width 2")

 

p3 <- p_data + geom_histogram(binwidth = 2, color = "grey30", 
                              fill = "red", alpha = 0.2) + 
  labs(title="Bin width 3")

 

grid.arrange(p1, p2, p3, ncol = 3)
```


### with `patchwork`

```{r}
p1 + p2 + p3
```

```{r}
p1 / (p2 +p3)
```


```{r}
p1 + p2 / p3
```

## Combining plots

```{r}
ggplot(mpg, aes(x = hwy)) + geom_histogram(aes(y=..density..), 
  color = "grey30", fill = "white") +
  geom_density(alpha = .2, fill = "antiquewhite3") + labs(title="Bin width 1")
```

```{r}
ggplot(mpg, aes(x = hwy)) + geom_histogram(aes(y=..density..), 
  color = "red", fill = "white") +
  geom_density(alpha = .2, fill = "antiquewhite3",
               color = "blue") +
  labs(title="Bin width 1")
```


```{r}
ggplot(mpg, aes(x = hwy)) + geom_histogram(aes(y=..density..), color = "grey30", fill = "white") +geom_density(alpha = .2, fill = "antiquewhite3") + labs(title="Bin width 1") + geom_vline(xintercept = mean(mpg$hwy), color = "red", linetype = "dashed")
```


```{r}
compare_mean <- mpg %>% 
  group_by(cyl) %>% 
  summarise(Mean = mean(cty))

g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill=factor(cyl)), 
                 alpha=0.5) +
    labs(title="Density plot",
         subtitle="City Mileage Grouped by Number of cylinders",
         caption="Source: mpg",
         x="City Mileage",
         fill="# Cylinders") +
  geom_vline(data = compare_mean,
             aes(xintercept = Mean, 
                 color = factor(cyl)),
             linetype = "dashed", size = 0.75)
```

## Error bars

```{r}
# Create new dataframe mpg_means_se
mpg_means_se <- mpg %>%
  group_by(manufacturer) %>% # Group the data by manufacturer
  summarize(mean_cty=mean(cty), # Create variable with mean of cty per group
            sd_cty=sd(cty), # Create variable with sd of cty per group
            N_cty=n(), # Create new variable N of cty per group
            se=sd_cty/sqrt(N_cty), # Create variable with se of cty per group
            upper_limit=mean_cty+se, # Upper limit
            lower_limit=mean_cty-se # Lower limit
            )
mpg_means_se
```

```{r}
ggplot(mpg_means_se, aes(x=manufacturer, y=mean_cty, fill=manufacturer)) +
  geom_errorbar(aes(ymin=lower_limit, ymax=upper_limit), width = 0.3) +
   geom_bar(stat="identity", alpha= 0.5) +
  theme(axis.text.x = element_text(angle=65, vjust=0.6), legend.position = "none") +
  geom_line(aes(group = 1), size = 0.2, color = "red")
```


```{r}
ggplot(mpg_means_se, aes(x=reorder(manufacturer, mean_cty), y=mean_cty, fill=manufacturer)) +
  geom_errorbar(aes(ymin=lower_limit, ymax=upper_limit), width = 0.3) +
   geom_bar(stat="identity", alpha= 0.5) +
  theme(axis.text.x = element_text(angle=65, vjust=0.6), legend.position = "none")# +
  #geom_line(aes(group = 1), size = 0.2, color = "red")
```

# Exercises

## `nycflights13`

Analyzing flights that departed from NYC in any of its airports.

```{r}
flights
```

### Histograms

```{r}
flights_ny <- flights %>% 
  unite(date, year:day, sep = "-") %>%
  mutate(date = ymd(date)) %>% 
  filter(dep_delay < 200) %>% 
  drop_na()
  
g <- flights_ny %>% 
  filter(dep_delay < 200) %>% 
  ggplot(aes(x = dep_delay))
ghist <- g +
  geom_histogram(bins = 100)

ghist + labs(title = "Departure delay")
```

```{r}
ghist + aes(fill = carrier)
```

```{r}
ghist + aes(fill = carrier) + facet_wrap(~ origin)
```

### Density plots

```{r}
gdens <- g +
  geom_density(aes(fill = carrier), alpha = 0.5) 
  
gdens + facet_wrap(~ origin)
```

```{r}
g <- flights_ny %>% 
  filter(month(date)==1,
         carrier == "AA") %>% 
  ggplot(aes(x = dep_time, y = dep_delay, 
             color = tailnum)) +
  geom_point(aes(frame = as.numeric(date))) + 
  facet_wrap(~ origin)
ggplotly(g)
```


```{r}
weather %>% 
  mutate(grupo = case_when(hour %in% 0:5 ~ "A",
                           hour %in% 6:11 ~ "B",
                           hour %in% 12:17 ~ "C",
                           TRUE ~ "D")
  )
```

