---
title: "Lecture 2"
subtitle: "`tidyverse` - `ggplot2` + `dplyr`"
author: "Pablo Benavides-Herrera"
date: 2020-05-30
output: 
  html_notebook:
    theme: spacelab
    toc: TRUE
    toc_float: TRUE
---

```{r global opts, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


We load the packages and take a look at the data `mpg`.

```{r pkgs}
library(tidyverse)
library(nycflights13)
mpg
```

# Using the `qplot()` function

```{r qplot}
qplot(displ, hwy, data = mpg)
```



```{r qplot + color}
qplot(displ, hwy, data = mpg, color = drv)
```

Now we change the `shape` of the points.

```{r qplot + color + shape}
qplot(displ, hwy, data = mpg, color = drv, shape = drv)
```

Now we add two geoms: points and smooth.

```{r qplot + geoms}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```

Smoothing across groups.

```{r qplot smooth by drv}
qplot(displ, hwy, data = mpg, color = drv, 
      geom = c("point", "smooth"))
```


```{r qplot hist}
qplot(hwy, data = mpg, fill = drv, binwidth = 2)
```

Plotting boxplots of `drv` and `hwy`.

```{r qplot boxplot}
qplot(drv, hwy, data = mpg, fill = drv, geom = "boxplot")
```

We can plot the histograms with facets.

```{r qplot + facets}
qplot(hwy, data = mpg, fill = drv, 
      facets = drv~., binwidth = 2)
```

Scatterplots can have facets too.

```{r qplot point + facet}
qplot(displ, hwy, data = mpg, facets = . ~ drv,
      geom = c("point","smooth"))
```

Getting the density plot.

```{r qplot density}
qplot(hwy, data = mpg, fill = drv,
      facets = . ~ drv,
      geom = "density")
```

# Translating `qplot()` into `ggplot()`

We'll do the same, but now using the `ggplot()` function.

The basic settings:

```{r ggplot config}
g <- ggplot(data = mpg)
g1 <- g + aes(x = hwy)
g2 <- g + aes(x = displ, y = hwy)
g3 <- g + aes(x = drv, y = hwy)
```


```{r ggplot}
g2 + geom_point()
```



```{r ggplot + color}
g2 + geom_point(aes(color = drv))
```

Now we change the `shape` of the points.

```{r ggplot + color + shape}
g2 + geom_point(aes(color = drv, shape = drv))
```

Now we add two geoms: points and smooth.

```{r ggplot + geoms}
g2 + geom_point() + geom_smooth()
```

Smoothing across groups.

```{r ggplot smooth by drv}
g2 + geom_point(aes(color = drv)) +
  geom_smooth(aes(color = drv))
```


```{r ggplot hist}
g1 + geom_histogram(aes(fill = drv), binwidth = 2)
```

Plotting boxplots of `drv` and `hwy`.

```{r ggplot boxplot}
g3 + geom_boxplot(aes(fill = drv))
```

We can plot the histograms with facets.

```{r ggplot + facets}
g1 + geom_histogram(aes(fill = drv)) +
  facet_wrap(~drv, ncol = 1)
```

Scatterplots can have facets too.

```{r ggplot point + facet}
g2 + geom_point() + geom_smooth() +
  facet_wrap(~ drv, nrow = 1)
```

Getting the density plot.

```{r ggplot density}
g1 + geom_density(aes(fill = drv)) +
  facet_wrap(~ drv, nrow = 1)
```

# Mouse Allergen and Asthma Cohort Study (MAACS)

“Are overweight individuals, as measured by body mass index (BMI), more susceptible than normal weight individuals to the harmful effects of PM2.5 on asthma symptoms?”

https://laqm.defra.gov.uk/public-health/pm25.html

NocturnalSymp, is the number of days in the past 2 weeks where the child experienced asthma symptoms (e.g. coughing, wheezing) while sleeping.

```{r}
maacs <- read_csv("bmi_pm25_no2_sim.csv")
glimpse(maacs)
```

```{r}
maacs
```

```{r}
g <- ggplot(data = maacs,
            aes(x = logpm25, y = NocturnalSympt))
summary(g)
```

With no geom, the plot is empty.

```{r}
print(g)
```

```{r}
g + geom_point()
```

```{r}
g + geom_point() + geom_smooth() 
```
Changing the smoothing method.

```{r}
g + geom_point() + geom_smooth(method = "lm") 
```

```{r}
g + geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~ bmicat)
```

```{r}
g + geom_point(aes(color = bmicat)) + 
  labs(title = "MAACS")
```

```{r}
g + geom_point(color = "steelblue", size = 4, alpha = 1/2)
```

```{r}
g + geom_point(aes(color = bmicat), size = 4, alpha = 1/2)
```

```{r}
g + geom_point(aes(color = bmicat)) +
  labs(title = "MAACS Cohort", 
       x = expression("log " * PM[2.5]), 
       y = "Nocturnal Symptoms")
```


```{r}
g + geom_point(aes(color = bmicat)) + 
  theme_bw(base_family = "Times")
```


https://ggplot2.tidyverse.org/reference/ggtheme.html
Complete themes — ggthemeThese are complete themes which control all non-data display. Use
theme() if you just need to tweak the display of an existing
theme.ggplot2.tidyverse.org​[9:53 a.m.] ALONSO SILVA, LAZARO



```{r}
testdat <- data.frame(x = 1:100, y = rnorm(100))
```


```{r}
testdat[50,2] <- 100  ## Outlier!
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()
```

```{r}
g + geom_line() + ylim(-3, 3)
```


```{r}
p <- g + geom_line() + coord_cartesian(ylim = c(-3, 3))
p
```

```{r}
p + annotate("text", x = c(45,55), y = c(2,2), label = c("here1", "here2"),
             color = c("red", "blue"))
```

# The `nycflights13` dataset

```{r}
flights
```

```{r}
daily <- flights %>% 
  group_by(dest, day) %>% 
  summarise(count = n())
daily
```

```{r}
daily %>% 
  ggplot(aes(x = day, y = count, color = dest)) +
  geom_point()
```

```{r}
daily %>% 
  ggplot(aes(x = day, y = count, color = dest)) +
  geom_line()
```

```{r}
daily %>% 
  distinct(dest)
```


```{r, fig.height= 8}
daily %>% 
  ggplot(aes(x = day, y = count, color = dest)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(trans = "log2") +
  geom_text(data = daily %>% ungroup() %>%  
              sample_n(size = 40), 
            aes(label = dest), color = "black") +
  theme(legend.key.size = unit(1.5, "cm"),
  legend.key.width = unit(0.5,"cm"))
```
```{r}
daily %>% 
  ungroup() %>% 
sample_n(40)
```





# `gapminder` dataset

```{r}
gap <- read_csv("gapminder-health-income.csv") %>% 
  mutate(population = log10(population))
gap
```

```{r}
gap %>% 
  ggplot(aes(x = income, y = health,
             color = population, 
             size = population)) +
  geom_point() +
  scale_color_viridis_c(
    guide = guide_colorbar(title.position = "right",
                          title.theme = element_text(angle = 90),
                          barheight = 20)) +
  scale_x_log10() +
  guides(size = FALSE) + 
  geom_text(data = gap %>% filter(country %in% 
                                    c("India",
                                      "Indonesia",
                                      "China",
                                      "United States",
                                      "Mexico")),
            aes(label = country), 
            color = "black", 
            size = 3,
            check_overlap = TRUE,
            hjust = 0,
            vjust = 1) +
  theme(legend.key.size = unit(1.5, "cm"),
  legend.key.width = unit(0.5,"cm"),
  legend.title.align = 0.5
  ) +
  labs(color = "Population (millions)")
```




